### 2014-11-21 wiml_poster/FCI_figure.tex
adds first draft of wiml_poster

\documentclass{article}
\usepackage{tikz} 
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx

\usepackage{tikz}

\usepackage{graphicx}
\usetikzlibrary{arrows,positioning} 
\usepackage{subcaption}
\thispagestyle{empty}
\usepackage{multicol,caption}
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}

\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}

\tikzset{
    %Define standard arrow tip
    >=stealth',
    %Define style for boxes
    observed/.style={
           circle,
           rounded corners,
           draw=black, thick,
           minimum width=.5em,
           minimum height=.5em,
           font=\footnotesize,
           text centered,
           fill=blue!20!white},
     latent/.style={
           circle,
           rounded corners,
           draw=black, thick, dashed,
           minimum width=.5em,
           minimum height=.5em,
           font=\footnotesize,
           text centered,
           fill=black!10!white
           },
    % Define arrow style
    pil/.style={
           o->,
           thick,
           shorten <=2pt,
           shorten >=2pt,},
    sh/.style={ shade, shading=axis, left color=red, right color=green,
    shading angle=45 }

}

\begin{document}
\def\ci{\perp\!\!\!\perp} % from Wikipedia

\begin{figure}
\centering
\begin{subfigure}[t]{0.44\textwidth}
\centering
\caption{}
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=0.45cm,  thick,main node/.style={observed}, shaded/.style={latent}]
\node[main node](1){C};
\node[main node, above left=of 1](2){A};
\node[main node, above right=of 1](3){B};
\node[main node, below=of 1](4){D};

 \path[]
 	(1) edge node {} (4)
    (2) edge [->]  (1)
    (3) edge [->]  (1);

\node[main node,right of=3,xshift=1.2em](5){A};
\node[main node, below right=of 5](6){C};
\node[main node, above right=of 6](7){B};
\node[main node, below=of 6](8){D};
\path[every node/.style={font=\sffamily\footnotesize}]
	(1) edge [-, >=stealth', blue!30!white, line width=.9em,shorten <=1em,shorten >= 1.7em,
	preaction={line width = .3em, blue!30!white,-triangle 90,draw,shorten >=.6em,shorten <=1em}
	] node[black]{FCI} (6)
	(5) edge [o->] (6)
	(7) edge [o->] (6)
	(6) edge [->] (8);

\end{tikzpicture}
\end{subfigure}
\begin{subfigure}[t]{0.55\textwidth}
\centering
\caption{}
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=0.45cm,  thick,main node/.style={observed}, shaded/.style={latent}]

\node[main node](1){A};
\node[main node, below right=of 1](2){B};
\node[shaded, above right=of 2](3){U};
\node[main node, below right=of 3](4){C};
\node[main node, above right=of 4](5){D};
\node[ below=of 3,yshift=.2cm](11){};
\path[] (1) edge (2)
		(3) edge (2) edge (4)
		(5) edge (4);

\node[main node,right=of 5,xshift=-.5em](6){A};
\node[main node, below right=of 6](7){B};
\node[above right=of 7](8){};
\node[main node, below right=of 8](9){C};
\node[main node, above right=of 9](10){D};
\node[below = of 8,yshift=0.2cm](12){};	
\path[every node/.style={font=\sffamily\footnotesize}]
(11) edge [-, >=stealth', blue!30!white, line width=.9em,shorten <=3.3em,shorten >= 4.5em,
	preaction={line width = .3em, blue!30!white,-triangle 90,draw,shorten >=3.7em,shorten <=4em}
	] node[black,xshift=-.5em]{FCI} (12)
(6) edge [o->] (7)
(7) edge [<->] (9)
(10) edge [o->](9);

\end{tikzpicture}
\end{subfigure}
\end{figure}

\end{document}