### 2015-01-15 causal-bandit/importance_sampling.tex

\documentclass{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{times}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage[ruled]{algorithm}
\usepackage{algpseudocode}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{amsthm}
\usepackage{natbib}
\usepackage[capitalize]{cleveref}
\usepackage{graphicx}
\usepackage{parskip}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}
\newcommand{\eqn}[1]{\begin{align}#1\end{align}}
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\Ber}{\operatorname{Bernoulli}}
\renewcommand{\P}[1]{\operatorname{P}\left\{#1\right\}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\title{Estimators}

\begin{document}
\def\ci{\perp\!\!\!\perp}
\maketitle

Suppose we have samples $(X_t,Y_t)$ for $t = 1... n$ drawn from the joint distribution $P(X,Y)$ and we want to estimate $\mu = P(Y=1|X=1)$

\section{The standard estimator}

\eqn{
\hat{\mu} = \frac{\sum_{t=1}^n Y_t \ind{X_t=1}}{\sum_{t=1}^n\ind{X_t=1}}
}

We can't actually say that this estimator is unbiased - as it can be undefined if $n = 0$

\color{red}{Is there an unbiased estimator for this simple problem if the $P(X = 1)$ is not known?}\color{black}

Let's consider variance and bounds.

\section{The importance sampling estimator}
If we know the probability that $X=1$ we can instead use importance sampling.

\eqn{
\hat{\mu} = \frac{1}{n}\sum_{t=1}^n\frac{Y_t \ind{X_t=1}}{P(X_t=1)}
}

The estimator is unbiased:

\eqn{
E[\hat{\mu}] = &\frac{1}{n}\sum_{t=1}^n E\left[\frac{Y_t \ind{X_t=1}}{P(X_t=1)}\right] = \frac{1}{np}\sum_{t=1}^n E\left[Y_t \ind{X_t=1}\right] \\
= & \frac{1}{n} \sum_{t=1}^n \frac{\left( 0*P(Y_t=0 \text{ or } X_t=0) + 1*P(Y_t = 1 \text{ and } X_t=1) \right)}{P(X_t=1)} \\
= & \frac{1}{n}\sum_{i=1}^n\frac{P(Y=1,X=1)}{P(X=1)} = \frac{1}{n}\sum_{i=1}^{n}P(Y=1|X=1) = \frac{1}{n} \sum_{i=1}^n \mu = \mu
}

We can use Hoeffding's inequality to get a bound on how far the estimator is likely to be from the true value. Let:

\eqn{
Z_t = \frac{Y_t \ind{X_t=1}}{p} \in \{0,\frac{1}{p}\}
}

\textbf{Hoeffding's inequality:} If $X_1...X_n$ are independent observations such that $a_i < X_i < b_i$ and $\bar{X} = \frac{1}{n}\sum_{i=1}^n X_i$ then:

\eqn {
P(\bar{X} - E[\bar{X}] \geq \epsilon) \leq \exp{\frac{-2n^2\epsilon^2}{\sum_{i=1}^n(b_i - a_i)^2}}
}

In our case this gives:
\eqn{
P(\frac{1}{n}\sum_{i=1}^n Z_t - \mu \geq \epsilon) \leq \exp{(-2n\epsilon^2p^2)} 
\implies P(\frac{1}{n}\sum_{i=1}^n Z_t - \mu \geq \frac{1}{p}\sqrt{\frac{1}{2n}\log{\frac{1}{\delta}}}) \leq \delta
}

This is not so good because $\frac{1}{p}$ can be very large if $p$ is small and its outside the log - so the bounds will grow quickly as $p$ gets small.

We can get a tighter bound by using an Chernoff's inequality that takes account of the variance of $X$. Let $W_t = \ind{Y_t = 1,X_t=1} = pZ_t$, then:

\eqn{
P(\frac{1}{n}\sum_{i=1}^n Z_t - \mu \geq \epsilon) = P(\frac{1}{n}\sum_{i=1}^n pZ_t - p\mu \geq p\epsilon) = P(\frac{1}{n}\sum_{i=1}^n W_t - p\mu \geq p\epsilon)
}

Now $W_t$ is a bernoulli random variable so:
\eqn{
V[W_t] = P(W_t=1)P(W_t=0) \leq P(W_t=1) = P(Y_t=1,X_t=1) \leq P(X_t=1) = p
}

\color{red} Now do the same for the standard estimator \color{black}

\bibliographystyle{plainnat}
\bibliography{c-bandit}

\end{document}