### 2015-11-13 causal-bandit/lowerbound_auer.tex
typed up auer lowerbound proof

\documentclass{article}

\usepackage{todonotes}

%\usepackage[disable]{todonotes}
\newcommand{\todot}[2][]{\todo[color=red!20!white,#1]{#2}}
\newcommand{\todof}[2][]{\todo[color=orange!20!white,#1]{#2}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% PACKAGES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{times}
\usepackage{fullpage}
\usepackage{latexsym}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[boxed]{algorithm}
\usepackage{algpseudocode}
\usepackage{mathtools}
\usepackage{accents}
\usepackage{tikz}
\usepackage{pgfplots}
\usepackage{dsfont}
\usepackage[bf]{caption}
\usepackage{hyperref}
\hypersetup{
    bookmarks=true,         % show bookmarks bar?
    unicode=false,          % non-Latin characters in AcrobatÕs bookmarks
    pdftoolbar=true,        % show AcrobatÕs toolbar?
    pdfmenubar=true,        % show AcrobatÕs menu?
    pdffitwindow=false,     % window fit to page when opened
    pdfstartview={FitH},    % fits the width of the page to the window
    pdftitle={My title},    % title
    pdfauthor={Author},     % author
    pdfsubject={Subject},   % subject of the document
    pdfcreator={Creator},   % creator of the document
    pdfproducer={Producer}, % producer of the document
    pdfkeywords={keyword1} {key2} {key3}, % list of keywords
    pdfnewwindow=true,      % links in new window
    colorlinks=true,       % false: boxed links; true: colored links
    linkcolor=red,          % color of internal links (change box color with linkbordercolor)
    citecolor=blue,        % color of links to bibliography
    filecolor=magenta,      % color of file links
    urlcolor=cyan           % color of external links
}
\usepackage{comment}
\usepackage{amsthm}
\usepackage{natbib}
\usepackage[capitalize]{cleveref}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{tikz} 
\usetikzlibrary{arrows,positioning} 
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}
\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MACROS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\defined}{\vcentcolon =}
\newcommand{\rdefined}{=\vcentcolon}
\newcommand{\E}{\mathbb E}
\newcommand{\Var}{\operatorname{Var}}
\newcommand{\calF}{\mathcal F}
\newcommand{\sr}[1]{\stackrel{#1}}
\newcommand{\set}[1]{\left\{#1\right\}}
\newcommand{\ind}[1]{\mathds{1}\!\!\set{#1}}
\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\newcommand{\floor}[1]{\left \lfloor {#1} \right\rfloor}
\newcommand{\ceil}[1]{\left \lceil {#1} \right\rceil}
\newcommand{\eqn}[1]{\begin{align}#1\end{align}}
\newcommand{\eq}[1]{\begin{align*}#1\end{align*}}
\newcommand{\Ber}{\operatorname{Bernoulli}}
\renewcommand{\P}[1]{\operatorname{P}\left\{#1\right\}}
\newcommand{\bigomega}[1]{\Omega\left( #1 \right)}

\renewcommand{\Pi}[1]{P_i\left( #1 \right)}
\newcommand{\Pu}[1]{P_{unif}\left( #1 \right)}
\newcommand{\Ps}[1]{P_{*}\left( #1 \right)}
\newcommand{\Ei}[1]{\E_i\left[ #1 \right]}
\newcommand{\Eu}[1]{\E_{unif}\left[ #1 \right]}
\newcommand{\Es}[1]{\E_{*}\left[ #1 \right]}
\renewcommand{\r}{\boldsymbol{r}}
\renewcommand{\log}[1]{lg\left( #1 \right)}
\newcommand{\kl}[2]{KL\left(#1 || #2 \right)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\begin{document}
\def\ci{\perp\!\!\!\perp}

\section{Lower bound on regret for K-armed bernoulli bandits}

For any horizon $T$ and number of arms $K$, there exists a strategy for choosing rewards such that the expected regret for any algorithm is $\bigomega{\sqrt{TK}}$. This strategy is oblivious to the algorithm and assigns rewards at random according to some distribution. Thus the regret bound applies to stochastic and adversarial bandits.

\begin{theorem}
There exits a distribution over rewards such that: 
\eqn {
R(T) \geq \frac{1}{20}\min{\{\sqrt{KT},T\}}
}
\end{theorem}

\begin{proof}

Consider a set of environments indexed by $i$. In environment $i$ action $i$ is the 'good' action. All other arms are slightly worse. 
\begin{itemize}
\item  In environment $i$, $r_{i,t} \sim Bernoulli(\frac{1}{2}+\epsilon)$ and $r_{j,t} \sim Bernoulli(\frac{1}{2}) \;\; \forall j \neq i$ 
\item $\Pi{.}$ is the probability with respect to environment $i$
\item $\Pu{.}$ is the probability with respect to an environment in which the expected reward for \textbf{all} arms is $\frac{1}{2}$
\item $\Ps{.}$ is the probability with respect to an environment sampled uniformly at random from $\set{1...K}$
\item $r_t$ is the reward received at time $t$
\item $\boldsymbol{r}^t = <r_1,...,r_t>$ is the history of rewards received upto time $t$
\item A is the algorithm, maps $\boldsymbol{r}^{t-1} \rightarrow i_t$
\item $N_i$ is a random variable representing the number of times action $i$ is selected by the algorithm.
\item $G_A = \sum_{t=1}^T r_t$ is the total reward for the algorithm
\item $G_{max}$ is the total reward for playing the optimal arm in every timestep.

\end{itemize}

\begin{lemma}
\label{lem:N_not_too_big}
For any arm $i$,
\eqn{
\Ei{N_i} -\Eu{N_i} \leq  \frac{T}{2}\sqrt{-\ln({1-4\epsilon^2})\Eu{N_i}}
}

This says that the number of times we expect to play arm $i$ in the environment in which it is optimal is not too much greater than the number of times we expect to play it if all arms are equal.

\begin{proof}
\eqn{
\Ei{N_i} -\Eu{N_i} = & \sum_{\r \in \set{0,1}^T}N_i(\r)\left(\Pi{\r} - \Pu{\r} \right) \qquad \leftarrow \text{definition of expectation}\\
\leq & \sum_{\r:\Pi{\r} \geq \Pu{\r}}N_i(\r)\left(\Pi{\r} - \Pu{\r} \right) \qquad \leftarrow \text{dropped only -ive terms}\\
\leq & T\sum_{\r:\Pi{\r} \geq \Pu{\r}}\left(\Pi{\r} - \Pu{\r} \right) \qquad \leftarrow N_i \leq T \;\; \forall \r\\
 = & \frac{T}{2}|| \Pi{\r} - \Pu{\r}||_1 \qquad \leftarrow \text{see Thomas\&Cover 11.137}\\
 \leq & \frac{T}{2}\sqrt{2\ln({2})\kl{\Pu{\r}}{\Pi{\r}}} \qquad \leftarrow \text{see Thomas\&Cover 11.138}\\
  = & \frac{T}{2}\sqrt{-ln(2)\log{1-4\epsilon^2} \Eu{N_i}} \qquad \leftarrow \text{see section \ref{sec:calculation_of_divergence}}\\
    = & \frac{T}{2}\sqrt{-\ln({1-4\epsilon^2}) \Eu{N_i}} \qquad \leftarrow \text{change of base}
}

\end{proof}
\end{lemma}

\begin{theorem}
For any algorithm $A$, if the distribution over rewards is selected unifirmly at random from environments $\set{1...K}$:

\eqn {
E_*[G_{max} - G_A] \geq \epsilon\left(T - \frac{T}{K} - \frac{T}{2}\sqrt{-\frac{T}{K}ln(1-4\epsilon^2)}  \right) 
}

\begin{proof}

\eqn{
\Ei{r_t} &= \left(\frac{1}{2}+\epsilon \right)\Pi{i_t = i} + \frac{1}{2}\left(1-\Pi{i_t = i} \right) \\
& = \frac{1}{2} + \epsilon \Pi{i_t = i}\\
\implies \Ei{G_A} &= \sum_{t=1}^T \Ei{r_t} = \frac{T}{2}+\epsilon \Ei{N_i}
}

This gives us the expected gain given action $i$ is the good action in terms of the number of times the algorithm $A$ selects action $i$. The expected gain over all the environments $i$ is:

\eqn{
\Es{G_A} &= \frac{1}{K}\sum_{i=1}^K \Ei{G_A} \\
&= \frac{T}{2} + \frac{\epsilon}{K}\sum_{i=1}^K \Ei{N_i}\\
\Es{G_{max}} &= \left(\frac{1}{2}+\epsilon \right)T
}

From lemma \ref{lem:N_not_too_big}

\eqn{
\sum_{i=1}^K\Ei{N_i} & \leq \sum_{i=1}^K \left(\Eu{N_i}+ \frac{T}{2}\sqrt{-\ln({1-4\epsilon^2})\Eu{N_i}}\right)
}

Now $\sum_{i=1}^K\Eu{N_i} = T$ (because $\sum_{i=1}^K N_i = T$ ? but doesn't that imply $\sum_{i=1}^K\Ei{N_i} = T$?)

\eqn{
&\implies \sum_{i=1}^K\Ei{N_i} \leq T+ \frac{T}{2} \sqrt{-\ln(1-4\epsilon^2)KT} \qquad \leftarrow \text{via Jenson's Inequality} \\
&\implies \Es{G_A} \leq \frac{T}{2} + \epsilon\left(\frac{T}{K}+ \frac{T}{2} \sqrt{-\frac{T}{K}\ln(1-4\epsilon^2)} \right) \\ 
&\implies \Es{G_{max} - G_A} \geq \epsilon\left(T - \frac{T}{K} - \frac{T}{2}\sqrt{-\frac{T}{K}ln(1-4\epsilon^2)}  \right) 
}

\end{proof}

\end{theorem}

\end{proof}

\subsection{Calculation of KL divergence}
\label{sec:calculation_of_divergence}
\eqn{
& \kl{p(x_1,...,x_n)}{q(x_1,...,x_n)} =  \sum_{i=1}^N \kl{p(x_i|\boldsymbol{x}^{i-1})}{q(x_i|\boldsymbol{x}^{i-1})} \qquad \leftarrow \text{chain rule for KL divergence} \\
& \text{where } \kl{p(x_i|\boldsymbol{x}^{i-1})}{q(x_i|\boldsymbol{x}^{i-1})} = 
\sum_{\boldsymbol{x}^{i-1}}p(\boldsymbol{x}^{i-1})\sum_{x_i}p(x_i|\boldsymbol{x}^{i-1})\log{\frac{p(x_i|\boldsymbol{x}^{i-1})}{q(x_i|\boldsymbol{x}^{i-1})}}
}

So 
\eqn {
\kl{P_{unif}}{P_i} = \sum_{t=1}^T\left(
\underbrace{\sum_{\r^{t-1}}\Pu{\r^{t-1}}}_{\text{expectation over history}}
\underbrace{\sum_{r_t \in \set{0,1}}\Pu{r_t|\r^{t-1}}\log{\frac{\Pu{r_t|\r^{t-1}}}{\Pi{r_t|\r^{t-1}}}}}_{\text{KL divergence at time t}}
\right)
}

Now 
\eqn{
\Pu{r_t|\r^{t-1}}& = \frac{1}{2} \;\; \forall \;{r_t,\r^{t-1}} \\
\Pi{r_t|\r^{t-1}}&=
\begin{cases}
(\frac{1}{2}+\epsilon)^{r_t}+(\frac{1}{2}-\epsilon)^{1-r_t} & \text{ if } A(\r^{t-1}) = i\\
\frac{1}{2} & \text{ otherwise}\\
\end{cases}
}

Let $B$ be the set of histories that lead the algorithm to select the good arm, $B =\set{ \r^{t-1}:A(\r^{t-1})=i}$

\eqn {
\kl{P_{unif}}{P_i} &= \sum_{t=1}^T\left(
\sum_{B}\Pu{\r^{t-1}}
\kl{\frac{1}{2}}{\frac{1}{2}+\epsilon}
+ \sum_{B^c}\Pu{\r^{t-1}}
\kl{\frac{1}{2}}{\frac{1}{2}}
\right)\\
&= \kl{\frac{1}{2}}{\frac{1}{2}+\epsilon}\sum_{t=1}^T\left(
\sum_{B}\Pu{\r^{t-1}}
\right)\\
&=\kl{\frac{1}{2}}{\frac{1}{2}+\epsilon} \sum_{t=1}^T\left(
\Pu{i_t=i}
\right)\\
&=\kl{\frac{1}{2}}{\frac{1}{2}+\epsilon}\Eu{N_i}\\
&=-\frac{1}{2}\log{1-4\epsilon^2}\Eu{N_i}
}
\pagebreak
\subsection{Jenson's inequality}
Jenson's inequality states that for a concave function $\phi$:

\eqn{
\frac{\sum_{i=1}^K \phi(x_i)}{K} \leq \phi \left(\frac{\sum_{i=1}^K x_i}{K} \right) \\
\implies  \sum_{i=1}^K \sqrt{\Eu{N_i}} \leq \sqrt{KT} \\
\implies \sum_{i=1}^K\Ei{N_i} \leq T+ \frac{T}{2} \sqrt{-\ln(1-4\epsilon^2)KT}
}

\end{document}