### 2015-12-04 causal-bandit/icml-2015/causal-bandits.tex
Start ICML draft

 %\documentclass[wcp,gray]{jmlr} % test grayscale version
\documentclass[wcp]{jmlr}

 % The following packages will be automatically loaded:
 % amsmath, amssymb, natbib, graphicx, url, algorithm2e

 %\usepackage{rotating}% for sideways figures and tables
\usepackage{longtable}% for long tables
\usepackage{booktabs}
% \usepackage[load-configurations=version-1]{siunitx} % newer version
 %\usepackage{siunitx}

% change the arguments, as appropriate, in the following:
\jmlrvolume{NNN}
\jmlryear{2015}
\jmlrworkshop{ICML}

\title[Causal Bandits]{Causal Bandits}

 % Use \Name{Author Name} to specify the name.
 % If the surname contains spaces, enclose the surname
 % in braces, e.g. \Name{John {Smith Jones}} similarly
 % if the name has a "von" part, e.g \Name{Jane {de Winter}}.
 % If the first letter in the forenames is a diacritic
 % enclose the diacritic in braces, e.g. \Name{{\'E}louise Smith}

 % Authors with different addresses:
 % \author{\Name{Author Name1} \Email{abc@sample.com}\\
 % \addr Address 1
 % \AND
 % \Name{Author Name2} \Email{xyz@sample.com}\\
 % \addr Address 2
 %}
\author{\Name{Author's Name}}

\editor{Editor's name}
 % \editors{List of editors' names}

\begin{document}

\maketitle

\begin{abstract}
	The abstract
\end{abstract}
\begin{keywords}
	Causality, Bandits, Regret Bounds
\end{keywords}

\section{Introduction}
\label{sec:intro}

The aim of this paper is to demonstrate theoretically the utility of having 
causal knowledge when solving multi-armed bandit (MAB) problems.

We do this by:
\begin{enumerate}
	\item Providing some motivating examples as to why it is reasonable to expect
		to have some knowledge of causal structures when solving a sequential
		decision making problem such as MAB problems.
	\item Formalising what we mean by ``causal information'' for a simple
		class of stochastic multi-armed bandit (MAB) problems. 
	\item Proposing a simple explore-then-exploit algorithm for this class of
		problems that makes use of the causal information and proving it
		has a sublinear regret bound.
	\item Deriving a matching lower bound for the same class of problems for 
		algorithms that make use of the causal structure.
	\item Showing that the upper bound for the ``causally informed'' algorithm
		is better than the corresponding lower bound for algorithms on this 
		class of problems that do not use the extra causal information.
\end{enumerate}

We identify a problem-dependent constant that appears in the upper and lower 
bounds that can be roughly interpreted as a measure of how much actions reveal
about other actions via the causal structure.

\subsection{Related Work}
\label{sub:related_work}

\begin{itemize}
	\item Bareinboim, Forney \& Pearl, \emph{Bandits with unobserved 
			confounders}, NIPS 2015
	\item Salomon, Audibert \& Alaoui, \emph{Lower Bounds and Selectivity of 
			Weak-Consistent Policies in Stochastic Multi-Armed Bandit Problems},
			JMLR 2013.
	\item Alon, Cesa-Bianchi, Dekel, Koren, \emph{Online Learning with Feedback
			Graphs: Beyond Bandits}, COLT 2015.
\end{itemize}

% \acks{Acknowledgements go here.}

\bibliography{../c-bandit.bib}

\end{document}