### 2017-10-03 thesis/talk/recipe_for_causal_inference.tex

\documentclass[tikz]{standalone}

\usepackage[latin1]{inputenc}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{xcolor}

\definecolor{talk}{HTML}{729FCF}
\definecolor{talk2}{HTML}{E8A753}
\definecolor{grey}{HTML}{DBDCDD}
\begin{document}
\pagestyle{empty}

\tikzset{
	%Define standard arrow tip
	>=stealth',
	%Define style for boxes
	punkt/.style={
		circle,
		rounded corners,
		draw=black, 
		text width=1em,
		minimum height=1em,
		text centered},
	observed/.style={
		circle,
		rounded corners,
		draw=black, thick,
		minimum width=2.3em,
		minimum height=2.3em,
		font=\footnotesize,
		text centered,
		fill=talk
	},
	latent/.style={
		circle,
		rounded corners,
		draw=black, thick, dashed,
		minimum width=2.2em,
		minimum height=2.2em,
		font=\footnotesize,
		text centered
	},
	target/.style={
		circle,
		rounded corners,
		draw=black, thick,
		minimum width=2.2em,
		minimum height=2.2em,
		font=\footnotesize,
		text centered,
		fill=black!20!white,
	},
	observedrect/.style={
		rectangle,
		rounded corners,
		draw=black, thick,
		minimum width=6em,
		minimum height=2em,
		font=\footnotesize,
		text centered,
		fill=talk!50
	},
	latentrect/.style={
		rectangle,
		rounded corners,
		draw=black, thick, dashed,
		minimum width=2.2em,
		minimum height=2.2em,
		font=\footnotesize,
		text centered
	},
	targetrect/.style={
		rectangle,
		rounded corners,
		draw=black, thick,
		minimum width=6em,
		minimum height=2em,
		font=\footnotesize,
		text centered,
		fill=black!20!white,
	},
	empty/.style={
		circle,
		rounded corners,
		minimum width=.5em,
		minimum height=.5em,
		font=\footnotesize,
		text centered,
	},
	% Define arrow style
	pil/.style={
		o->,
		thick,
		shorten <=2pt,
		shorten >=2pt,},
	sh/.style={ shade, shading=axis, left color=red, right color=green,
		shading angle=45 },
	line/.style={
		draw,
		-latex
	},
    ln/.style={
    	draw
    },
	block/.style={
		rectangle, 
		draw, 
		fill=talk!50, 
		text width=8em, 
		text centered, 
		rounded corners, 
		minimum height=3em
	}   
}

\begin{tikzpicture}[node distance = 2cm, node/.style={punkt}]
% Place nodes
\node[node](or){or};
\node[block, right of = or](learngraph){Apply \emph{Causal Discovery} to learn the graph};
\node [block,left of=or] (drawgraph) {Use prior knowledge to draw the graph};
\node [block, below of=or] (docalc) {Apply the Do Calculus};
\node [block, below left of=docalc,node distance=3.5cm] (fail) {Non Identifiable};
\node [block, below right of=docalc, node distance=3.5cm] (success) {Identifiable};
\node [block, below of=fail](notident){Bugger! You will have to make additional assumptions or do experiments};
\node [block, below of=success](ident){Great! Apply the formula from the do-caclulus to estimate effect.};
\node [block, fill=talk2!50, dashed, right of=ident,node distance=4.5cm](stats){Deal with pesky statistical issues introduced by finite data as you see fit};

% Draw edges
\path[ln](or) -- (learngraph);
\path[ln](or) -- (drawgraph);
\path[line] (or) -- (docalc);
\path[line] (docalc) -- (fail);
\path[line] (docalc) -- (success);
\path[line] (fail) -- (notident);
\path[line] (success) -- (ident);
\path[line] (stats) -- (ident);

\end{tikzpicture}

\end{document}