### 2017-10-14 thesis/talk/do_calculus_examples.tex

\documentclass{article}
\usepackage{tikz} 
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{listings}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{tabularx}
\usepackage{enumitem}
\usepackage{algorithm}% http://ctan.org/pkg/algorithm
\usepackage[noend]{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{standalone}

\usepackage{subcaption}
\usepackage{multicol,caption}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={210mm,297mm},
 left=20mm,
 right=20mm,
 top=20mm,
 bottom=20mm,
 }

\usetikzlibrary{arrows,positioning,shapes} 
\pgfarrowsdeclarecombine{ring}{ring}{}{}{o}{o}
\thispagestyle{empty}
\DeclareMathOperator{\ringarrow}{\raisebox{0.5ex}{\tikz[baseline]{\draw[ring->](0,0)--(2em,0);}}}

\definecolor{talk}{HTML}{729FCF}
\definecolor{talk2}{HTML}{E8A753}
\definecolor{grey}{HTML}{DBDCDD}
\renewcommand{\vec}[1]{\boldsymbol{#1}}
\tikzset{
    %Define standard arrow tip
    >=stealth',
    %Define style for boxes
    observed/.style={
   	circle,
   	rounded corners,
   	draw=black, thick,
   	minimum width=2.3em,
   	minimum height=2.3em,
   	font=\footnotesize,
   	text centered,
   	fill=talk!50
   },
   latent/.style={
   	circle,
   	rounded corners,
   	draw=black, thick, dashed,
   	minimum width=2.2em,
   	minimum height=2.2em,
   	font=\footnotesize,
   	text centered
   },
    % Define arrow style
    pil/.style={
           o->,
           thick,
           shorten <=2pt,
           shorten >=2pt,},
    sh/.style={ shade, shading=axis, left color=red, right color=green,
    shading angle=45 },
observedrect/.style={
	rectangle,
	rounded corners,
	draw=black, thick,
	minimum width=3em,
	minimum height=1.5em,
	font=\footnotesize,
	text centered,
	fill=talk!50
},  
 target/.style={
	circle,
	rounded corners,
	draw=black, thick,
	minimum width=2.3em,
	minimum height=2.3em,
	font=\footnotesize,
	text centered,
	fill=talk!50
},
empty/.style={
	circle,
	rounded corners,
	minimum width=.5em,
	minimum height=.5em,
	font=\footnotesize,
	text centered,
},
}

\begin{document}
\pagestyle{empty}

\begin{figure}
	\centering
	\begin{subfigure}[t]{0.3\textwidth}
	\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=5em,thick,node/.style={observed}, tg/.style={target},act/.style={observed},lt/.style={empty}]
	\node[lt](3){};
	\node[act, below left of=3](1){X};
	\node[tg, below right of = 3](2){Y};
	\node[above=of 1,yshift=-4em](T){\includegraphics[width=1cm]{tick.png}};
	\path[]
	(1) edge (2);
	\end{tikzpicture}
	\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=5em,thick,node/.style={observed}, tg/.style={target},act/.style={observed},lt/.style={latent}]
	\node[lt](3){U};
	\node[act, below left of=3](1){X};
	\node[tg, below right of = 3](2){Y};
	\node[above=of 1,yshift=-4em](T){\includegraphics[width=1cm]{cross.png}};
	\path[]
	(1) edge (2)
	(3) edge (1) edge (2);
	\end{tikzpicture}
\end{subfigure}
\begin{subfigure}[t]{0.3\textwidth}
	\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=5em,thick,node/.style={observed}, tg/.style={target},act/.style={observed},lt/.style={latent}]
	\node[lt](3){U};
	\node[act, below left of=3](1){X};
	\node[tg, below right of = 3](2){Y};
	\node[node,above left of=1](4){Z};
	\node[above=of 1,yshift=-4em](T){\includegraphics[width=1cm]{cross.png}};
	\path[]
	(1) edge (2)
	(3) edge (1) edge (2)
	(4) edge (1);
	\end{tikzpicture}
\end{subfigure}
	\begin{subfigure}[t]{0.3\textwidth}
	\vspace{3em}
	\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=5em,thick,node/.style={observed}, tg/.style={target},act/.style={observed},lt/.style={empty}]
\node[node](3){Z};
\node[act, below left of=3](1){X};
\node[tg, below right of = 3](2){Y};
\node[above=of 1,yshift=-4em](T){\includegraphics[width=1cm]{tick.png}};
\path[]
(1) edge (2)
(3) edge (1) edge (2);
\end{tikzpicture}
\end{subfigure}
\begin{subfigure}[t]{0.3\textwidth}
	\vspace{3em}
	\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=5em,thick,node/.style={observed}, tg/.style={target},act/.style={observed},lt/.style={latent}]
	\node[act](1){X};
	\node[node,right of = 1](4){Z};
	\node[lt, above of=4,yshift=-1.5em](3){U};
	\node[tg, right of = 4](2){Y};
	\node[above=of 1,yshift=-4em](T){\includegraphics[width=1cm]{tick.png}};
	\path[]
	(1) edge (4)
	(4) edge (2)
	(3) edge (1) edge (2);
	\end{tikzpicture}
\end{subfigure}
\begin{subfigure}[t]{0.3\textwidth}
	\vspace{3em}
\begin{tikzpicture}[->,shorten >=0pt,shorten <=0pt,node distance=5em,thick,node/.style={observed}, tg/.style={target},act/.style={observed},lt/.style={latent}]
\node[act](1){X};
\node[node,right of = 1](4){Z};
\node[lt, above of=4,yshift=-1.5em](3){U};
\node[tg, right of = 4](2){Y};
\node[lt, below right of = 4,xshift=-.9em,yshift=0.5em](5){U};
\node[above=of 1,yshift=-4em](T){\includegraphics[width=1cm]{cross.png}};
\path[]
(1) edge (4)
(4) edge (2)
(3) edge (1) edge (2)
(5) edge (4) edge (2);
\end{tikzpicture}
\end{subfigure}
\end{figure}

\end{document}